---
title: "Exploring Education Data Portal"
author: "Crystal Lewis"
date: "10/10/2021"
output: html_document
---

<br>

This is very bare bones data exploration into the Urban Institute's Education Data Portal. I reviewed enrollment trends for the 3 high schools in Columbia Public Schools, Columbia, MO. This is the town where I went to high school.

```{r, echo = FALSE, warning = FALSE, message = FALSE}

library(logr)
library(tidyverse)
library(educationdata)
library(flextable)

```


Use the API to pull enrollment data. Then bind data together.

```{r, message = FALSE, warning = FALSE, evaluate = FALSE}

# Rockbridge

enroll_rock <- get_education_data(level = 'schools', 
                         source = 'ccd', 
                         topic = 'enrollment', 
                         filters = list(year=2015:2019, 
                                        ncessch = '290100000299'))

# Hickman

enroll_hick <- get_education_data(level = 'schools', 
                         source = 'ccd', 
                         topic = 'enrollment', 
                         filters = list(year=2015:2019, 
                                        ncessch = '290100000300'))

# Battle

enroll_bat <- get_education_data(level = 'schools', 
                         source = 'ccd', 
                         topic = 'enrollment', 
                         filters = list(year=2015:2019, 
                                        ncessch = '290100003239'))

# Directory (for school name)

dir <- get_education_data(level = 'schools', 
                         source = 'ccd', 
                         topic = 'directory', 
                         filters = list(year=2019,
                                        leaid = '2901000'))

# Bind 3 schools together and filter to just total enrollment (grade 99 = total)

# dim(enroll_bat)
# dim(enroll_hick)
# dim(enroll_rock)

enroll <- bind_rows(enroll_bat, enroll_hick, enroll_rock) %>%
  filter (grade == 99)

# str(enroll)
# skimr::skim(enroll)

# Join enrollment with directory for school name

dir_small <- dir %>%
  select(ncessch, school_name)

enroll_names <- enroll %>%
  left_join(dir_small, by = "ncessch")

# dim(enroll_names)


```


Visualize enrollment over time

```{r, fig.align="center"}

enroll_names %>%
  ggplot(aes(x=year, y=enrollment, color=school_name)) +
  geom_line(stat = "identity")+
  geom_point() + 
  labs(title = "Enrollment by school by year", color = "School Name") +
  scale_color_manual(values = c("sienna1", "hotpink1", "hotpink4")) +
  theme_classic()
```

Produce a table of enrollment numbers over time

```{r}

enroll_names %>% 
  select(school_name, year, enrollment) %>%
  pivot_wider(names_from = year, 
              values_from = enrollment,
              names_prefix = "enroll_") %>% 
  relocate(school_name, enroll_2015, enroll_2016, enroll_2017, enroll_2018, enroll_2019) %>%
  set_names(toupper) %>%
  rename_with(~ str_replace(., "_", " ")) %>%
  flextable() %>%
  add_header_lines(values = "Columbia Public Schools Enrollment 2015-2019", top = TRUE)


```


